---
title: "Week 4 – Statistics: Distributions & First Model"
author: "Group Name / Members"
date: "`r format(Sys.Date())`"
format:
  html:
    toc: false
    toc-depth: 2
---

## Overview {.unnumbered}

::: callout-note
### What this file is for

This week is hands-on with **statistics**. You will:
- Measure spread (**SD, IQR**) for two numerical variables
- Visualise **distribution shapes** (histograms + Normal overlays)
- Fit a **simple linear regression** (e.g., `grade ~ study_hours`)
- Inspect **residuals** to check model fit
- Record brief **group reflections**

Work with your **cleaned Week 2 dataset** (no impossible or extreme values).
:::

---

# Dataset & Setup

Use your cleaned dataset prepared in Week 2. Mind the file path.

```{r}
library(tidyverse)

# Load your cleaned dataset


# Quick structure check

```

**Confirm (1–2 lines):** that the dataset loaded correctly (rows/columns) and that `grade` and `study_hours` are numeric.

---

# Spread & Variability

Compute summaries, SD, and IQRs for at least two two main variables from your dataset (e.g., `grade` and `study_hours`).

```{r}
# Summaries



```

**Comment (2–3 lines):** Which variable varies more? Are the ranges reasonable? What does the **IQR** say about the middle 50%?

---

# Distribution Shapes

Plot histograms and compare with a Normal curve for the same variables.

```{r}
# Histogram and normal curve for variable 1





```

**Comment (2–3 lines):** Is each variable roughly Normal, skewed, or heavy-tailed? Does the red curve fit well? Any gaps or outliers?

---

# First Model (Simple Linear Regression)

Fit a simple linear model (use appropriate variables and reason your choice).

```{r}
# create a model usinf lm and provide summary


```

**Interpretation (3–4 lines):**
- Intercept (β₀) = ___ means …
- Slope (β₁) = ___ means …
- (Optional) Is the slope statistically significant (p-value)?
- (Optional) How much variation does the model explain (R²)?

::: callout-note
### What is a model?

A **model** is a simplified, mathematical explanation of how things relate or behave.  
It is **expressed mathematically** (equation/algorithm), **summarises** relationships, **simplifies** what matters, **generalises (within limits)**, and **must be interpretable**.

> A good model is clear, honest, and useful — not a perfect copy of reality, but a way to think better about it.
:::

---

# Model Fit (Residuals)

Residuals are **errors** = observed − predicted. They should look like random noise around 0.

```{r}
# Plot the residuals including abline


```

**Comment (2–3 lines):** Are residuals roughly centred around 0 and randomly scattered? Any patterns or extreme points? What does this suggest about model fit?

---

# Group Reflection

Keep answers short and specific (1–2 sentences each).

| Question | Reflection |
|:--|:--|
| **Distributions:** What did shapes reveal about variation or consistency? | |
| **Model:** What does your regression line suggest about the relationship? | |
| **Learning:** What was the main thing you learned about your dataset this week? | |

---

# Expected Outputs

Your file should include (at least):
- Two distribution plots
- One scatterplot with fitted regression line
- One residual plot
- Regression summary with brief interpretation
- Completed reflection table

---

# You are free to explore further along modeling (e.g., trying different variable combinations, or trying multiple regression, or different models available in R libraries).
# You may also explore along p-values and R².
