---
title: "FDS Group Presentation"
date: "`r format(Sys.Date())`"
format: 
  revealjs:
    theme: serif
    slide-number: true
    title-slide-attributes:
      data-background-color: "#f4f4ed"
    footer: "Group_36_FDS_Group_Presentation"
editor: visual
---

```{r}
#| include: false 
# use of Gen AI:
# I have used Gen AI , model GPT5 to 

# - generate many ideas for different research areas on the particular topics given in the course. 
# Ex. Prompt: Could you generate ideas of study in climate change, Energy use,Carbon emissions, Renewable energy sources, Environmental policy from our world in data for data analysis studies?

# - Once we formed the question and hypothesis, we checked the data avalibality in the website our world in data using AI
# - Ex. Prompt: Could you check data availability for annual temperature, Co2 emission.. etc from our world in data website.

# - After our study , some the results required deeper knowledge of the overall climate dynamics or overall european  climate poclies?, i have asked articles that can be coraborated with my results to find why it behaves that way( I have references those articles used)
# Ex prompt- My results show that rho = 0.32, which is not very high, could you let me know why it behaves this way through some articles?
# Ex prompt- WHy does many countries see shift in energy intensity and fossil fuel share after 1990? what policy changes has happened after that?

# - I have had no experience in html format coding, i need to know how to change fonts, colors, placement of pictures, tables, text in a particular fit. So i used AI to generate syntax for these HTML presentation formatting.
# Ex prompt - Could let me know how to place the table right of a pciture using html formatting?
# Ex prompt - how ot add captions for a pciture using html formatting?
# and also https://quarto.org/docs/presentations/revealjs/advanced.html which has been sited.

```

<h3 style="font-size: 2.7em; margin-top: 0;text-align: center;">

Linking Carbon Emissions to Climate Change

</h3>

<hr style="width: 600px; margin: 50px auto; border-top: 5px solid #333;">

<h3 style="font-size: 1.2em; margin-bottom: 5px;font-style: italic;text-align: center;">

Investigating the CO₂ Drivers and

</h3>

<h3 style="font-size: 1.2em; margin-top: 0;font-style: italic;text-align: center;">

Temperature Effects in the Netherlands

</h3>

```{=html}
<style type="text/css">
/* Only this slide */
.reveal section.toc-slide h2 { font-size: 1.4em; }
.reveal section.toc-slide h3 { font-size: 1.1em; margin: 0.1em 0; }
.reveal section.toc-slide h4 { font-size: 0.9em; margin: 0.1em 0; }
.reveal section.toc-slide p  { font-size: 0.65em;  margin: 0.1em 0 0.6em 0; }
</style>
```

## Table of Contents {.toc-slide}

<hr style="width: 1000px; margin: 10px 0; border-top: 5px solid #333;">

:::::: columns
::: {.column width="33%"}
<h2 style="font-size: 2.0em; margin-top: 0; font-style: bold; text-align: center;">

01

</h2>

<h3 style="font-size: 1.0em; margin-top: 0; font-style: bold; text-align: center;">

Introduction

</h3>

<h4 style="font-size: 0.5em; margin-top: 0; font-style: italic; text-align: center;">

Setting the scene!

</h4>

<br>

<h2 style="font-size: 2.0em; margin-top: 0; font-style: bold; text-align: center;">

04

</h2>

<h3 style="font-size: 1.0em; margin-top: 0; font-style: bold; text-align: center;">

Methodology

</h3>

<h4 style="font-size: 0.5em; margin-top: 0; font-style: italic; text-align: center;">

From data to insight!

</h4>
:::

::: {.column width="33%"}
<h2 style="font-size: 2.0em; margin-top: 0; font-style: bold; text-align: center;">

02

</h2>

<h3 style="font-size: 1.0em; margin-top: 0; font-style: bold; text-align: center;">

Statement

</h3>

<h4 style="font-size: 0.5em; margin-top: 0; font-style: italic; text-align: center;">

What we set out to know?

</h4>

<br>

<h2 style="font-size: 2.0em; margin-top: 0; font-style: bold; text-align: center;">

05

</h2>

<h3 style="font-size: 1.0em; margin-top: 0; font-style: bold; text-align: center;">

Analysis

</h3>

<h4 style="font-size: 0.5em; margin-top: 0; font-style: italic; text-align: center;">

Examining signals up close!

</h4>
:::

::: {.column width="33%"}
<h2 style="font-size: 2.0em; margin-top: 0; font-style: bold; text-align: center;">

03

</h2>

<h3 style="font-size: 1.0em; margin-top: 0; font-style: bold; text-align: center;">

Hypotheses

</h3>

<h4 style="font-size: 0.5em; margin-top: 0; font-style: italic; text-align: center;">

Expectations before evidence!

</h4>

<br>

<h2 style="font-size: 2.0em; margin-top: 0; font-style: bold; text-align: center;">

06

</h2>

<h3 style="font-size: 1.0em; margin-top: 0; font-style: bold; text-align: center;">

Conclusion

</h3>

<h4 style="font-size: 0.5em; margin-top: 0; font-style: italic; text-align: center;">

Final takeaway insights!

</h4>
:::
::::::

## Introduction {.toc-slide}

<hr style="width: 1000px; margin: 10px 0; border-top: 5px solid #333;">

:::::: columns
::: {.column width="33%"}
#### What is climate change?

Refers to long-term shifts in temperature and weather patterns.
:::

::: {.column width="33%"}
#### Who is responsible?

![*Fig 1: Global GHG emissions by sector – 2016 (1)*](public.avif){width="100%" fig-align="left"}
:::

::: {.column width="33%"}
#### What is carbon footprint?

Total GHG caused by a person, organization, product, or activity, expressed as CO₂ equivalent.
:::
::::::

<h3 style="font-size: 1.0em; margin-top: 0; font-style: italic; text-align: left;">

Is excess greenhouse gas (GHG) emissions, especially carbon dioxide (CO₂) the central driver for climate change?

</h3>

## 02 Research Questions {.toc-slide}

<hr style="width: 1000px; margin: 10px 0; border-top: 5px solid #333;">

### 📝 Statements

<h3 style="font-size: 1.6em; margin-top:100px 0;text-align: left;">

Statement 1.

</h3>

<h4 style="font-size: 1.3em; margin-top: 0;text-align: left;">

*How strongly are regional, for ex. in Netherlands, annual temperature increase is correlated with annual CO₂ emissions, and how has this relationship changed over the last few years?*

<h4>

<hr style="width: 500px; margin: 50px auto; border-top: 5px solid #333;">

<h3 style="font-size: 1.6em; margin-top:100px 0;text-align: left;">

Statement 2.

</h3>

<h4 style="font-size: 1.3em; margin-top: 0;text-align: left;">

*What are the largest contributors to recent changes in CO₂ emissions \~ energy efficiency, GDP growth, or energy mix for Netherlands?*

<h4>

## 03 Hypotheses

<hr style="width: 1000px; margin: 10px 0; border-top: 5px solid #333;">

#### 💡 Hypotheses for statement 1. (RQ1)

<h3 style="font-size: 0.9em; margin-top: 0;text-align: left;">

Q1.2: NL Temperature Trend

<h3>

<h4 style="font-size: 0.8em; margin-top: 0;text-align: left;font-weight: normal">

**Hypothesis:** The linear trend (slope) of Netherlands annual mean temperature over time is not zero.

<h4>

<h3 style="font-size: 0.9em; margin-top: 0;text-align: left;">

Q1.2: NL-Global Co-variance

<h3>

<h4 style="font-size: 0.8em; margin-top: 0;text-align: left;font-weight: normal">

**Hypothesis:** There is linear association between detrended NL temperature anomalies and detrended global temperature anomalies.

<h4>

<h3 style="font-size: 0.9em; margin-top: 0;text-align: left;">

Q1.3: Global CO₂/Temperature Association

<h3>

<h4 style="font-size: 0.8em; margin-top: 0;text-align: left;font-weight: normal">

**Hypothesis:** Changes in atmospheric CO₂ are associated with changes in global mean surface temperature.

<h4>

## 03 Hypotheses

<hr style="width: 1000px; margin: 10px 0; border-top: 5px solid #333;">

#### 💡 Hypotheses for statement 2. (RQ2)

<h3 style="font-size: 0.9em; margin-top: 0;text-align: left;">

Q2.1: NL\~CO₂ drivers Multi-variant regression analysis

<h3>

<h4 style="font-size: 0.75em; margin-top: 0;text-align: left;font-weight: normal">

**Hypothesis:** GDP growth is the strongest and most consistent driver of annual CO₂ emission changes in the Netherlands, even when controlling for fossil-fuel share and energy-intensity changes.

<h4>

<h3 style="font-size: 0.6em; margin-top: 0;text-align: left;">

Regression Model

</h3>

<p style="font-size: 0.6em; margin-top: 0;text-align: left;">

Δln(CO₂<sub>t</sub>) = β<sub>0</sub> + β<sub>1</sub> Δln(GDP<sub>t</sub>) + β<sub>2</sub> Δln(FossilShare<sub>t</sub>) + β<sub>3</sub> Δln(EnergyIntensity<sub>t</sub>) + ε<sub>t</sub>

</p>

<h3 style="font-size: 0.7em; margin-top: 0;text-align: left;">

Where:

</h3>

<ul style="font-size: 0.5em; margin-top: 0;text-align: left;">

<li><b>Δln(CO₂<sub>t</sub>)</b> = annual growth rate of CO₂ emissions</li>

<li><b>Δln(GDP<sub>t</sub>)</b> = annual GDP growth</li>

<li><b>Δln(FossilShare<sub>t</sub>)</b> = percent change in fossil-fuel share</li>

<li><b>Δln(EnergyIntensity<sub>t</sub>)</b> = percent change in energy intensity</li>

<li><b>ε<sub>t</sub></b> = error term</li>

</ul>

## 04 Methodolgy

<hr style="width: 1000px; margin: 10px 0; border-top: 5px solid #333;">

| [DATA SOURCES]{style="font-size: 0.55em;"} | [$\longrightarrow$]{style="font-size: 0.55em;"} | [DATA PROCESSING]{style="font-size: 0.55em;"} | [$\longrightarrow$]{style="font-size: 0.55em;"} | [STATISTICAL MODELS]{style="font-size: 0.55em;"} | [$\longrightarrow$]{style="font-size: 0.55em;"} | [PERIOD COMPARISON]{style="font-size: 0.5em;"} | [$\longrightarrow$]{style="font-size: 0.5em;"} | [INSIGHTS]{style="font-size: 0.5em;"} |   |
|:------:|:------:|:------:|:------:|:------:|:------:|:------:|:------:|:------:|:------:|
| [*Action:* Temps(3), CO₂(4),GDP(5)(6), Energy(7)(8)(9)(10)]{style="font-size: 0.55em;"} |  | [*Method:* Cleaning, Transforming, detrending, Growth rates (Δln)]{style="font-size: 0.55em;"} |  | [*Method:*<br>Trends, covariance, Regression analyses, Multi-variant regression]{style="font-size: 0.55em;"} |  | [*Method:*<br> Pre vs Post 1990 Shifts in drivers]{style="font-size: 0.55em;"} |  | [Drivers, impacts, Final storyline]{style="font-size: 0.55em;"} |  |

## 04 Analysis

<hr style="width: 1000px; margin: 10px 0; border-top: 5px solid #333;">

#### Q1.1: Does Netherlands annual mean temperature over time follow a linear trend?

```{r}
#| include: false  

library(tidyverse)
library(lubridate)
library(dplyr)

# Step 1: Load the Dataset 

# Fetch the data
Annual_temperature <- read.csv("https://ourworldindata.org/grapher/average-monthly-surface-temperature.csv?v=1&csvType=filtered&useColumnShortNames=true&tab=line&time=earliest..2025-10-15&country=NLD~Europe+%28NIAID%29~OWID_WRL&overlay=download-data")
head(Annual_temperature, 10)

# Fetch the data carbon emissions_per capita
Carbon_emission <- read.csv("https://ourworldindata.org/grapher/co-emissions-per-capita.csv?v=1&csvType=filtered&useColumnShortNames=true&country=OWID_WRL~OWID_EU27~NLD&overlay=download-data")

head(Carbon_emission, 10)

# Fetch the data carbon emission total
Carbon_emission_tl <- read.csv("https://ourworldindata.org/grapher/annual-co2-emissions-per-country.csv?v=1&csvType=filtered&useColumnShortNames=true&country=OWID_WRL~OWID_EUR~NLD&overlay=download-data")
head(Carbon_emission_tl, 10)
```

```{r}
#| include: false  

# Step 2 : Initial inspection
glimpse(Annual_temperature)
summary(Annual_temperature)

glimpse(Carbon_emission)
summary(Carbon_emission)

glimpse(Carbon_emission_tl)
summary(Carbon_emission_tl)
```

```{r}
#| include: false  
# Step 3 : Data cleaning
# 3.1.1 : Count missing values in each column
colSums(is.na(Annual_temperature))

# 3.1.2: Remove rows with any missing values
Annual_temperature_clean <- na.omit(Annual_temperature)

# 3.1.3 : formating
Annual_temperature_clean$entity <- str_trim(Annual_temperature_clean$entity )

# 3.2.1 : Count missing values in each column
colSums(is.na(Carbon_emission))

#  3.2.2 : Remove rows with any missing values
Carbon_emission_clean <- na.omit(Carbon_emission)

#  3.2.3  : formating
Carbon_emission$entity <- str_trim(Carbon_emission$entity )

# 3.3.1 : Count missing values in each column
colSums(is.na(Carbon_emission_tl))

#  3.3.2 : Remove rows with any missing values
Carbon_emission_tl_clean <- na.omit(Carbon_emission_tl)

#  3.3.3  : formating
Carbon_emission_tl$entity <- str_trim(Carbon_emission_tl$entity )
names(Carbon_emission)
names(Carbon_emission_tl)

```

```{r}
#| include: false  
# Step 4: Transformation, integration, validation
Annual_temperature_group <- Annual_temperature_clean |> 
  mutate(year = year(ymd(day)))|>
  group_by(year, entity) |> 
  summarise (Avg_temp = round ( mean (temperature_2m),2))

NLD_AT <- Annual_temperature_group |>
  filter (
    entity == "Netherlands",          # Only Netherlands countries
    year >= 1940 & year <= 2024 # Year range
  )
EU_AT <- Annual_temperature_group |>
  filter (
    entity == "Europe (NIAID)",          # Overall Europe
    year >= 1940 & year <= 2024 # Year range
  )
WD_AT <- Annual_temperature_group |>
  filter (
    entity == "World",          # Overall Europe
    year >= 1940 & year <= 2024 # Year range
  )
```

```{r}
#| include: false  
# Step 5 : Descriptive statistics
summary(NLD_AT)
summary(EU_AT)
summary(WD_AT)
```

::::: columns
::: {.column width="68%"}
```{r}
# Step 6.1:  Visualization 
# Linear trend plot 
base_size <- 18
ggplot(Annual_temperature_group, aes(x = year,
                                     y = Avg_temp,
                                     color = entity, group = entity)) +
  geom_line(size = 0.5) +
  geom_point(size = 0.5) +
  geom_smooth(method = "lm",
              formula = y ~ poly(x, 2, raw = TRUE),
              aes(color = entity),     # <-- IMPORTANT: separate lines,
              se = TRUE) +

  theme_minimal(base_size = base_size) + 
  theme(
    plot.background  = element_rect(fill = "#f4f4ed", colour = NA),   
    panel.background = element_rect(fill = "#f4f4ed", colour = NA),
    panel.grid.major = element_line(colour = "#dddddd"),
    plot.title = element_text(size = 22, face = "bold",colour = "#333333"),
    axis.text.x = element_text(size = 16,angle = 90, hjust = 1,colour = "#333333"),
    axis.text.y = element_text(size = 16,colour = "#333333"),
    axis.title.y = element_text(size = 16,colour = "#333333"),
    axis.title = element_text(face = "bold",colour = "#333333"),
    legend.text = element_text(size = 16,colour = "#333333"),
    legend.title = element_text(size = 18,colour = "#333333")
  ) +

 labs(
    title = "Annual average temperature  (Netherlands Vs Europe & World)",
    y = "Annual average temperature(°C) ",
    x = "",
    color = "Region"
  ) +
  scale_x_continuous(
    breaks = seq(1940, 2024, 5),
    labels = seq(1940, 2024, 5)
  )
```
:::

::: {.column width="32%"}
<h2 style="font-size: 0.58em; text-align: left;">

Statistical Model Fit Summary

</h2>

| [Region]{style="font-size: 0.5em;"} | [$\mathbf{R}^2$]{style="font-size: 0.5em;"} | [$\mathbf{Pr(>|t|)}$]{style="font-size: 0.5em;"} |
|:--------------------|:------------------------:|:------------------------:|
| [**World**]{style="font-size: 0.5em;"} | [0.9217]{style="font-size: 0.5em;"} | [2.70e-16]{style="font-size: 0.5em;"} |
| [**Europe**]{style="font-size: 0.5em;"} | [0.6349]{style="font-size: 0.5em;"} | [0.000264]{style="font-size: 0.5em;"} |
| [**Netherlands**]{style="font-size: 0.5em;"} | [0.4941]{style="font-size: 0.5em;"} | [0.00131]{style="font-size: 0.5em;"} |
:::
:::::

<h2 style="font-size: 0.6em; text-align: left;">

*-Netherlands temperatures show a significant linear warming trend since the 1970s, tracking and slightly exceeding European/global rates.*<br> *- Moderate R² values reflect high inter annual variability from internal climate dynamics (11).*

</h2>

## 04 Analysis

<hr style="width: 1000px; margin: 10px 0; border-top: 5px solid #333;">

#### Q1.2: Detrended Covariance \~ NL and Global Temp.

::::: columns
::: {.column width="68%"}
```{r}
# Step 6.2:  Visualization 
# Detrend the time series using linear model residuals

WD_Temp_scatter <-  full_join(WD_AT, NLD_AT, by = "year")
WD_Temp_scatter_LM <- WD_Temp_scatter |>
   mutate(Nld_temp = Avg_temp.y, Wd_temp = Avg_temp.x)|>
   select(-Avg_temp.x,-Avg_temp.y)
WD_Temp_scatter_LM <-WD_Temp_scatter_LM[, c(1, 2, 5, 3, 4)]

nl_detrended <- resid(lm(WD_Temp_scatter_LM$Nld_temp ~ seq_along(WD_Temp_scatter_LM$Nld_temp)))
global_detrended <- resid(lm(WD_Temp_scatter_LM$Wd_temp ~ seq_along(WD_Temp_scatter_LM$Wd_temp)))

plot_df <- data.frame(
  year = NLD_AT$year,
  NL = nl_detrended,
  Global = global_detrended
)

# Plot both series
ggplot(plot_df, aes(x = Global, y = NL)) +
  geom_point(color = "#333333", alpha = 0.6) +
  geom_smooth(method = "lm", color = "#333333", se = TRUE) +
  labs(
    title = "Scatter Plot of Detrended - NL vs Global Temperature Anomalies",
    x = "Global Detrended Temperature(°C)",
    y = "NL Detrended Temperature (°C)"
  ) +
  theme_minimal() + 
  theme(
    plot.margin = margin(t = 10, r = 10, b = 15, l = 15),
    plot.background  = element_rect(fill = "#f4f4ed", colour = NA),   
    panel.background = element_rect(fill = "#f4f4ed", colour = NA),
    panel.grid.major = element_line(colour = "#dddddd"),
    plot.title = element_text(size = 21, face = "bold",colour = "#333333",margin = margin(b = 8)),
    axis.text.x = element_text(size = 16,colour = "#333333"),
    axis.text.y = element_text(size = 16,colour = "#333333"),
    axis.title.y = element_text(size = 16,colour = "#333333"),
    axis.title.x = element_text(size = 16,colour = "#333333"),
    axis.title = element_text(face = "bold",colour = "#333333"),
  ) +
  scale_x_continuous(
    breaks = seq(-0.5, 0.5, 0.10),
    labels = seq(-0.5, 0.5, 0.10)
  ) 
```
:::

::: {.column width="32%"}
<h2 style="font-size: 0.58em; text-align: left;">

Statistical Model Fit Summary

</h2>

| [Region]{style="font-size: 0.5em;"} | [$\mathbf{\rho}$]{style="font-size: 0.5em;"} | [$\mathbf{Pr(>|t|)}$]{style="font-size: 0.5em;"} |
|:--------------------|:------------------------:|:------------------------:|
| [**NLD/GBL**]{style="font-size: 0.5em;"} | [0.3242]{style="font-size: 0.5em;"} | [0.00247]{style="font-size: 0.5em;"} |
:::
:::::

<h2 style="font-size: 0.60em; text-align: left;">

*-Detrended Dutch (NL) temperatures show a modest but significant positive co-variation with global temperature anomalies (ρ ≈ 0.32).*<br><br> *-The relationship is statistically significant (p ≈ 0.002), indicating synchronized variability beyond long-term trends(12).*

</h2>

## 04 Analysis

<hr style="width: 1000px; margin: 10px 0; border-top: 5px solid #333;">

#### Q1.3: Global CO₂ Concentration vs Temp. Relationship

::::: columns
::: {.column width="68%"}
```{r}
# Step 6.3:  Visualization 
# regression plot
WD_CE <- Carbon_emission_tl_clean |>
  filter(year >= 1940 & entity == "World")|>
  rename(year = year)

WD_scatter <-   full_join(WD_AT, WD_CE, by = "year")
WD_scatter <- WD_scatter|>
  filter(year >= 1940)

ggplot(WD_scatter, aes(x = emissions_total/100000000, y = Avg_temp)) +
    geom_point(aes(),size = 0.75, color = "#333333") +
    geom_smooth(method = "lm",size = 0.75, color = "#333333", se = TRUE) +
    theme_minimal() +
    theme(
      plot.subtitle = element_text(size = 16),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_text(size = 16),
      axis.title.x = element_text(size = 16),
      axis.title.y = element_text(size = 16),
      plot.margin = margin(t = 10, r = 10, b = 15, l = 15),
      plot.background  = element_rect(fill = "#f4f4ed", colour = NA),   
      panel.background = element_rect(fill = "#f4f4ed", colour = NA),
      panel.grid.major = element_line(colour = "#dddddd"),
      plot.title = element_text(size = 21, face = "bold",colour = "#333333",margin = margin(b = 8)),
      text = element_text(color = "#333333"),
      axis.text = element_text(color = "#333333"),
      axis.title = element_text(color = "#333333", face = "bold")
    ) +
   labs(
     title = "Average temperature vs CO₂ emissions (Global)",
     y = "Annual average temperature (°C) ",
     x = "Annual CO₂ emissions (billion-tons)"
   ) 
```
:::

::: {.column width="32%"}
<h2 style="font-size: 0.58em; text-align: left;">

Statistical Model Fit Summary

</h2>

| [Region]{style="font-size: 0.5em;"} | [$\mathbf{R}^2$]{style="font-size: 0.5em;"} | [$\mathbf{Pr(>|t|)}$]{style="font-size: 0.5em;"} |
|:--------------------|:------------------------:|:------------------------:|
| [**World**]{style="font-size: 0.5em;"} | [0.8401]{style="font-size: 0.5em;"} | [2.2e-16]{style="font-size: 0.5em;"} |
:::
:::::

<h2 style="font-size: 0.60em; text-align: left;">

*-Global CO₂ levels show a very strong statistical association with temperature increases (R² = 0.84), consistent with the established physical link.*<br><br> *-The results clearly indicate that increasing CO₂ emissions are the primary driver of warming globally and in the Netherlands.*

</h2>

## 04 Analysis

<hr style="width: 1000px; margin: 10px 0; border-top: 5px solid #333;">

#### Key CO₂ drivers \~ fossil fuel share and energy efficency?

```{r}
#| include: false 
# Step 1: Load the Dataset 
# Fetch the data
GDP_per_capita <- read.csv("https://ourworldindata.org/grapher/gdp-per-capita-maddison-project-database.csv?v=1&csvType=filtered&useColumnShortNames=true&tab=line&time=1965..latest&country=NLD~OWID_WRL&overlay=download-data")
head(GDP_per_capita, 10)
SPE_FF <- read.csv("https://ourworldindata.org/grapher/fossil-fuels-share-energy.csv?v=1&csvType=filtered&useColumnShortNames=true&tab=line&country=NLD~OWID_WRL~OWID_EUR&overlay=download-data")
head(SPE_FF, 10)
SPE_TE <- read.csv("https://ourworldindata.org/grapher/primary-sub-energy-source.csv?v=1&csvType=filtered&useColumnShortNames=true&country=~NLD&overlay=download-data")
head(SPE_TE, 10)
SPE_EU_TE <- read.csv("https://ourworldindata.org/grapher/primary-sub-energy-source.csv?v=1&csvType=filtered&useColumnShortNames=true&country=~OWID_EUR&overlay=download-data")
head(SPE_EU_TE, 10)
FGUK_GDP_per_captia <- read.csv("https://ourworldindata.org/grapher/gdp-per-capita-maddison-project-database.csv?v=1&csvType=filtered&useColumnShortNames=true&tab=line&time=1963..latest&country=GBR~DEU~FRA&overlay=download-data")
head(FGUK_GDP_per_captia)
UK_SPE_TE <- read.csv("https://ourworldindata.org/grapher/primary-sub-energy-source.csv?v=1&csvType=filtered&useColumnShortNames=true&country=~GBR&overlay=download-data")
GER_SPE_TE <- read.csv("https://ourworldindata.org/grapher/primary-sub-energy-source.csv?v=1&csvType=filtered&useColumnShortNames=true&country=~DEU&overlay=download-data")
FRA_SPE_TE <- read.csv("https://ourworldindata.org/grapher/primary-sub-energy-source.csv?v=1&csvType=filtered&useColumnShortNames=true&country=~FRA&overlay=download-data")
```

```{r}
#| include: false 
# Step 2: Initial inspection
glimpse(GDP_per_capita)
summary(GDP_per_capita)

glimpse(SPE_FF)
summary(SPE_FF)

glimpse(SPE_TE)
summary(SPE_TE)

glimpse(FGUK_GDP_per_captia)
summary(FGUK_GDP_per_captia) 
```

```{r}
#| include: false 
# Step 3: Data cleaning
# 3.1.1 : Count missing values in each column
colSums(is.na(GDP_per_capita))

# 3.1.2 : Remove ows with any missing values
GDP_per_capita_clean <- GDP_per_capita
  #select(-X900793.annotations )|>
  #na.omit()

# 3.1.3 : formating
GDP_per_capita_clean$Entity <- str_trim(GDP_per_capita_clean$entity)

# 3.2.1 : Count missing values in each column
colSums(is.na(SPE_FF))

# 3.2.2 : Remove rows with any missing values
SPE_FF_clean <- SPE_FF|>
  na.omit()

# 3.2.3 : formating
SPE_FF_clean$entity <- str_trim(SPE_FF_clean$entity)

# 3.3.1 : Count missing values in each column
colSums(is.na(SPE_TE))

# 3.3.3 : Remove rows with any missing values
SPE_TE_clean <- SPE_TE|>
  na.omit()

# 3.3.4 : formating
SPE_TE_clean$entity <- str_trim(SPE_TE_clean$entity)

# 3.4.1 : Count missing values in each column
colSums(is.na(FGUK_GDP_per_captia))

# 3.4.2 : Remove rows with any missing values
FGUK_GDP_per_captia_clean <- FGUK_GDP_per_captia
  #select(-X900793.annotations )|>
  #na.omit()

# 3.4.3 : formating
FGUK_GDP_per_captia_clean$entity <- str_trim(FGUK_GDP_per_captia_clean$entity)

# 3.5.1 : Count missing values in each column for UK total energy
colSums(is.na(UK_SPE_TE))

# 3.5.2 : Remove rows with any missing values
UK_SPE_TE_clean <- UK_SPE_TE |>
  na.omit()

# 3.5.3 : formating
UK_SPE_TE_clean$entity <- str_trim(UK_SPE_TE_clean$entity)

# 3.6.1 : Count missing values in each column for Germany total energy
colSums(is.na(GER_SPE_TE))

# 3.6.2 : Remove rows with any missing values
GER_SPE_TE_clean <- GER_SPE_TE |>
  na.omit()

# 3.6.3 : formating
GER_SPE_TE_clean$entity <- str_trim(GER_SPE_TE_clean$entity)

# 3.7.1 : Count missing values in each column for France total energy
colSums(is.na(FRA_SPE_TE))

# 3.7.2 : Remove rows with any missing values
FRA_SPE_TE_clean <- FRA_SPE_TE |>
  na.omit()

# 3.7.3 : formating
FRA_SPE_TE_clean$entity <- str_trim(FRA_SPE_TE_clean$entity)
```

```{r}
#| include: false 
# Step 4: Transformation, integration, validation
NLD_GDP <- GDP_per_capita_clean |>
  filter (
    entity == "Netherlands",          # Only Netherlands countries
    year >= 1965 & year <= 2022 # Year range
  )
EU_GDP <- GDP_per_capita_clean |>
  filter (
    entity == "Europe (UN)",          # Overall Europe
    year >= 1954 & year <= 2022 # Year range
  )
WD_GDP <- GDP_per_capita_clean|>
  filter (
    entity == "World",          # Overall Europe
    year >= 1965 & year <= 2022 # Year range
  )

NLD_SPE_FF <- SPE_FF_clean |>
  filter (
    entity == "Netherlands",          # Only Netherlands countries
    year >= 1965 & year <= 2022 # Year range
  )
EU_SPE_FF<- SPE_FF_clean  |>
  filter (
    entity == "Europe (UN)",          # Overall Europe
    year >= 1954 & year <= 2022 # Year range
  )
WD_SPE_FF <- SPE_FF_clean |>
  filter (
    entity == "World",          # Overall Europe
    year >= 1965 & year <= 2022 # Year range
  )

NLD_SPE_TE <- SPE_TE_clean |>
  filter (
    entity == "Netherlands",          # Only Netherlands countries
    year >= 1965 & year <= 2022)|> # Year range
  mutate(TE = rowSums(across(c(4, 5,6,7,8,9,10,11,12))))

NLD_CE <- Carbon_emission_tl_clean |>
  filter(
    year >= 1965 & year <= 2022,
    entity == "Netherlands")

NLD_Analysis2 <- NLD_GDP|>
  left_join(NLD_SPE_FF, by = "year")|>
  left_join(NLD_SPE_TE, by = "year")|>
  left_join(NLD_CE, by = "year")|>
  select(entity.x, code.x, year, gdp_per_capita, fossil_fuels__pct_equivalent_primary_energy, TE, emissions_total)|>
  mutate(Energy_intensity = TE/gdp_per_capita)
    
 NLD_Analysis_pre <- NLD_Analysis2 |>
  filter(
    year >= 1965 & year <=1989)

NLD_Analysis_post <- NLD_Analysis2 |>
   filter(
    year >= 1990 & year <= 2022)
summary(NLD_Analysis2)

# combine UK GER FRA 
EU3_SPE_TE <- bind_rows (FRA_SPE_TE_clean,GER_SPE_TE_clean,UK_SPE_TE_clean)
head(EU3_SPE_TE)

EU3_SPE_TE <- EU3_SPE_TE|>
  filter (
    year >= 1965 & year <= 2022)|> # Year range
  mutate(TE = rowSums(across(c(4,5,6,7,8,9,10,11,12))))|> 
  mutate(entity = str_trim(entity))

EU3_Analysis2 <- EU3_SPE_TE|>
  left_join(FGUK_GDP_per_captia_clean, by = c("year","entity"))|>
  select(entity, code.x, year, gdp_per_capita, TE)|>
  mutate(Energy_intensity = TE/gdp_per_capita)

summary(EU3_Analysis2) 
```

::::: columns
::: {.column width="50%"}
<p>[Percentage of total primary energy consumption supplied by coal, oil, and gas.]{style="font-size: 0.5em;"}</p>

```{r}
# Step 4.1:  Pre-Visualization 
plot_spe_ff <- SPE_FF_clean |>
  filter(entity %in% c("Netherlands", "Europe", "World")) |>
  filter(year >= 1965 & year <= 2022)

ggplot(plot_spe_ff,
       aes(x = year,
           y = fossil_fuels__pct_equivalent_primary_energy,
           color = entity,
           group = entity)) +
  geom_line(size = 0.8,) +
  geom_point(size = 1) +
  scale_color_manual(values = c("Netherlands" = "#333333", "Europe" = "#2E8B57", World = "#32CD32")) +
  theme_minimal(base_size = 14) +
  labs(
    title = "Share of primary energy from fossil fuels (NLD vs EU vs WD)",
    x = " ",
    y = "Fossil fuels share of energy (%)",
    color = "Region"
  ) +
  scale_x_continuous(
    breaks = seq(1965, 2022, 5),
    labels = seq(1965, 2022, 5)
  ) +
  theme(
      plot.subtitle = element_text(size = 16),
      axis.text.x = element_text(size = 16,angle = 90, hjust =1),
      axis.text.y = element_text(size = 16),
      axis.title.x = element_text(size = 16),
      axis.title.y = element_text(size = 16),
      plot.margin = margin(t = 10, r = 10, b = 15, l = 15),
      plot.background  = element_rect(fill = "#f4f4ed", colour = NA),   
      panel.background = element_rect(fill = "#f4f4ed", colour = NA),
      panel.grid.major = element_line(colour = "#dddddd"),
      plot.title = element_text(size = 21, face = "bold",colour = "#333333",margin = margin(b = 8)),
      text = element_text(color = "#333333"),
      axis.text = element_text(color = "#333333"),
      axis.title = element_text(color = "#333333", face = "bold"),
      legend.text = element_text(size = 16,colour = "#333333"),
      legend.title = element_text(size = 18,colour = "#333333")
  )
```
:::

::: {.column width="50%"}
<p>[Energy Intensity = (Total Energy / GDP), <br> where, total energy = sum of primary energy sources]{style="font-size: 0.5em;"}</p>

```{r}
# Step 4.2:  Pre-Visualization 
# Netherlands Vs top 3 EU countries
# Netherlands data for plotting
EI_NLD <- NLD_Analysis2 |>
  mutate(Region = "Netherlands") |>
  select(Region, year, Energy_intensity)

#  (UK, Germany, France) for plotting
EI_EU3 <- EU3_Analysis2 |>
  mutate(Region = entity) |>
  select(Region, year, Energy_intensity)

# Bind into one plotting data frame
EI_plot_df <- bind_rows(EI_NLD, EI_EU3)

ggplot(EI_plot_df,
       aes(x = year,
           y = Energy_intensity,
           color = Region,
           group = Region)) +
  geom_line(size = 0.8) +
  scale_color_manual(values = c("Netherlands" = "#333333", "France" = "#2E8B57", Germany = "#32CD32", "United kingdom" = "#98FB98")) +
  geom_point(size = 1) +
  theme_minimal(base_size = 14) +
  labs(
    title = "Energy intensity (primary energy per unit of GDP)",
    x = " ",
    y = "Energy intensity (total energy / GDP)",
    color = "Region"
  ) +
  scale_x_continuous(
    breaks = seq(min(EI_plot_df$year), max(EI_plot_df$year), 5)
  ) +
  theme(
      plot.subtitle = element_text(size = 16),
      axis.text.x = element_text(size = 16,angle = 90, hjust =1),
      axis.text.y = element_text(size = 16),
      axis.title.x = element_text(size = 16),
      axis.title.y = element_text(size = 16),
      plot.margin = margin(t = 10, r = 10, b = 15, l = 15),
      plot.background  = element_rect(fill = "#f4f4ed", colour = NA),   
      panel.background = element_rect(fill = "#f4f4ed", colour = NA),
      panel.grid.major = element_line(colour = "#dddddd"),
      plot.title = element_text(size = 21, face = "bold",colour = "#333333",margin = margin(b = 8)),
      text = element_text(color = "#333333"),
      axis.text = element_text(color = "#333333"),
      axis.title = element_text(color = "#333333", face = "bold"),
      legend.text = element_text(size = 16,colour = "#333333"),
      legend.title = element_text(size = 18,colour = "#333333")
  )

```
:::
:::::

<h2 style="font-size: 0.5em; text-align: left;">

*-Fossil fuel shares fall - decline becomes much steeper after about 1990, especially for the Netherlands relative to Europe and the world.*<br> *-Energy intensity drops - improves much faster after about 1990, with all becoming more efficiency driven.*<br> *-So, we decided to split our regression analysis into two periods to study these effects(pre and post 1990)*

</h2>

## 04 Analysis

<hr style="width: 1000px; margin: 10px 0; border-top: 5px solid #333;">

#### Q2.1 NL\~CO₂ drivers - GDP growth

```{r}
#| include: false 
# Step 5.1 : Modeling 
# Multi-variant regression analysis pre 1990
# log differences for growth rates
NLD_Analysis_pre$CO2Growth <- c(NA, diff(log(NLD_Analysis_pre$emissions_total)))
NLD_Analysis_pre$GDPGrowth <- c(NA, diff(log(NLD_Analysis_pre$gdp_per_capita)))
NLD_Analysis_pre$EIGrowth <- c(NA, diff(log(NLD_Analysis_pre$Energy_intensity)))
NLD_Analysis_pre$FSGrowth <- c(NA, diff(log(NLD_Analysis_pre$fossil_fuels__pct_equivalent_primary_energy)))

# regression model
Annova_model <- lm(CO2Growth ~ GDPGrowth + FSGrowth + EIGrowth, data = NLD_Analysis_pre)

#  regression summary
summary(Annova_model)
```

```{r}
#| include: false 
# Step 5.1 : Modeling 
# Multi-variant regression analysis post 1990
# log differences for growth rates
NLD_Analysis_post$CO2Growth <- c(NA, diff(log(NLD_Analysis_post$emissions_total)))
NLD_Analysis_post$GDPGrowth <- c(NA, diff(log(NLD_Analysis_post$gdp_per_capita)))
NLD_Analysis_post$EIGrowth <- c(NA, diff(log(NLD_Analysis_post$Energy_intensity)))
NLD_Analysis_post$FSGrowth <- c(NA, diff(log(NLD_Analysis_post$fossil_fuels__pct_equivalent_primary_energy)))

# regression model
Annova_90_model <- lm(CO2Growth ~ GDPGrowth + FSGrowth + EIGrowth, data = NLD_Analysis_post)

# regression summary
summary(Annova_90_model)
```

::::: columns
::: {.column width="68%"}
```{r}
# Step 6.1:  Visualization 
NLD_Analysis_pre <- NLD_Analysis_pre |> mutate(period = "Pre-1990")
NLD_Analysis_post <- NLD_Analysis_post |> mutate(period = "Post-1990")
NLD_Analysis_plot <- bind_rows(NLD_Analysis_pre, NLD_Analysis_post)

# 1. GDPGrowth vs CO2Growth 
ggplot(NLD_Analysis_plot, aes(x = GDPGrowth, y = CO2Growth, color = period)) +
  geom_point(size = 2, alpha = 0.7,color = "#333333") +
  scale_color_manual(values = c("Pre-1990" = "#333333", "Post-1990" = "#2E8B57")) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "CO₂ Growth vs GDP Growth (Pre vs Post 1990)",
       x = "GDP Growth (log difference)",
       y = "CO₂ Growth (log difference)",
       color = "Period") +
  theme_minimal(base_size = 14)+
theme(
      plot.subtitle = element_text(size = 16),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_text(size = 16),
      axis.title.x = element_text(size = 16),
      axis.title.y = element_text(size = 16),
      plot.margin = margin(t = 10, r = 10, b = 15, l = 15),
      plot.background  = element_rect(fill = "#f4f4ed", colour = NA),   
      panel.background = element_rect(fill = "#f4f4ed", colour = NA),
      panel.grid.major = element_line(colour = "#dddddd"),
      plot.title = element_text(size = 21, face = "bold",colour = "#333333",margin = margin(b = 8)),
      text = element_text(color = "#333333"),
      axis.text = element_text(color = "#333333"),
      axis.title = element_text(color = "#333333", face = "bold")
    ) 


```
:::

::: {.column width="32%"}
<h2 style="font-size: 0.58em; text-align: left;">

Statistical Model Fit Summary

</h2>

| [Region]{style="font-size: 0.5em;"} | [$\mathbf{\beta}$]{style="font-size: 0.5em;"} | [$\mathbf{Pr(>|t|)}$]{style="font-size: 0.5em;"} |
|:--------------------|:------------------------:|:------------------------:|
| [**Pre-1990**]{style="font-size: 0.5em;"} | [2.14119]{style="font-size: 0.5em;"} | [0.00778\*]{style="font-size: 0.5em;"} |
| [**Post-1990**]{style="font-size: 0.5em;"} | [0.966597]{style="font-size: 0.5em;"} | [4.3e-05\*]{style="font-size: 0.5em;"} |
:::
:::::

<h2 style="font-size: 0.65em; text-align: left;">

*-strong, statistically significant driver of CO₂ growth in both periods.* <br><br> *-The effect size drops after 1990, suggesting economies became somewhat less carbon-intensive even as growth continued to raise emissions.*

<h2>

## 04 Analysis

<hr style="width: 1000px; margin: 10px 0; border-top: 5px solid #333;">

#### Q2.1 NL\~CO₂ drivers - Fossil Fuel Share growth

::::: columns
::: {.column width="68%"}
```{r}
# Step 6.2:  Visualization 
# 2. FSGrowth vs CO2Growth
ggplot(NLD_Analysis_plot, aes(x = FSGrowth, y = CO2Growth, color = period)) +
  geom_point(size = 2, alpha = 0.7, color = "#333333") +
  geom_smooth(method = "lm", se = TRUE) +
  scale_color_manual(values = c("Pre-1990" = "#333333", "Post-1990" = "#2E8B57")) +
  labs(title = "CO₂ Growth vs Fossil Share Growth (Pre vs Post 1990)",
       x = "Fossil Share Growth (log difference)",
       y = "CO₂ Growth (log difference)",
       color = "Period") +
  theme_minimal(base_size = 14) +
theme(
      plot.subtitle = element_text(size = 16),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_text(size = 16),
      axis.title.x = element_text(size = 16),
      axis.title.y = element_text(size = 16),
      plot.margin = margin(t = 10, r = 10, b = 15, l = 15),
      plot.background  = element_rect(fill = "#f4f4ed", colour = NA),   
      panel.background = element_rect(fill = "#f4f4ed", colour = NA),
      panel.grid.major = element_line(colour = "#dddddd"),
      plot.title = element_text(size = 21, face = "bold",colour = "#333333",margin = margin(b = 8)),
      text = element_text(color = "#333333"),
      axis.text = element_text(color = "#333333"),
      axis.title = element_text(color = "#333333", face = "bold")
    ) 
```
:::

::: {.column width="32%"}
<h2 style="font-size: 0.58em; text-align: left;">

Statistical Model Fit Summary

</h2>

| [Region]{style="font-size: 0.5em;"} | [$\mathbf{\beta}$]{style="font-size: 0.5em;"} | [$\mathbf{Pr(>|t|)}$]{style="font-size: 0.5em;"} |
|:--------------------|:------------------------:|:------------------------:|
| [**Pre-1990**]{style="font-size: 0.5em;"} | [4.03462]{style="font-size: 0.5em;"} | [0.51390]{style="font-size: 0.5em;"} |
| [**Post-1990**]{style="font-size: 0.5em;"} | [1.399934]{style="font-size: 0.5em;"} | [0.00353\*]{style="font-size: 0.5em;"} |
:::
:::::

<h2 style="font-size: 0.65em; text-align: left;">

*-CO₂ growth shows a weak, non-significant link to fossil-fuel share changes before 1990, but a strong, significant positive relationship emerges after 1990.*<br><br> *-This suggests fossil-fuel dependency became a much clearer driver of emissions growth in the post-1990 period.*

<h2>

## 04 Analysis

<hr style="width: 1000px; margin: 10px 0; border-top: 5px solid #333;">

#### Q2.1 NL\~CO₂ drivers - Energy intensity growth

::::: columns
::: {.column width="68%"}
```{r}
# Step 6.3:  Visualization 
# 3. EIGrowth vs CO2Growth
ggplot(NLD_Analysis_plot, aes(x = EIGrowth, y = CO2Growth, color = period)) +
  geom_point(size = 2, alpha = 0.7, color = "#333333") +
  scale_color_manual(values = c("Pre-1990" = "#333333", "Post-1990" = "#2E8B57")) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "CO₂ Growth vs Energy Intensity Growth (Pre vs Post 1990)",
       x = "Energy Intensity Growth (log difference)",
       y = "CO₂ Growth (log difference)",
       color = "Period") +
  theme_minimal(base_size = 14) +
theme(
      plot.subtitle = element_text(size = 16),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_text(size = 16),
      axis.title.x = element_text(size = 16),
      axis.title.y = element_text(size = 16),
      plot.margin = margin(t = 10, r = 10, b = 15, l = 15),
      plot.background  = element_rect(fill = "#f4f4ed", colour = NA),   
      panel.background = element_rect(fill = "#f4f4ed", colour = NA),
      panel.grid.major = element_line(colour = "#dddddd"),
      plot.title = element_text(size = 21, face = "bold",colour = "#333333",margin = margin(b = 8)),
      text = element_text(color = "#333333"),
      axis.text = element_text(color = "#333333"),
      axis.title = element_text(color = "#333333", face = "bold")
    ) 
```
:::

::: {.column width="32%"}
<h2 style="font-size: 0.58em; text-align: left;">

Statistical Model Fit Summary

</h2>

| [Region]{style="font-size: 0.5em;"} | [$\mathbf{\beta}$]{style="font-size: 0.5em;"} | [$\mathbf{Pr(>|t|)}$]{style="font-size: 0.5em;"} |
|:--------------------|:------------------------:|:------------------------:|
| [**Pre-1990**]{style="font-size: 0.5em;"} | [0.67796]{style="font-size: 0.5em;"} | [0.02726]{style="font-size: 0.5em;"} |
| [**Post-1990**]{style="font-size: 0.5em;"} | [0.861386]{style="font-size: 0.5em;"} | [6.2e-07\*]{style="font-size: 0.5em;"} |
:::
:::::

<h2 style="font-size: 0.65em; text-align: left;">

*-Strong and significant driver of CO₂ growth both before and after 1990.* <br> <br> *-The relationship intensifies post-1990, indicating that efficiency improvements matter even more for emissions outcomes in the modern period.*

<h2>

## Conclusion– Key Takeaways

<hr style="width: 1000px; margin: 10px 0; border-top: 5px solid #333;">

<h2 style="font-size: 0.8em; text-align: left;">

RQ1 findings:<br> - Netherlands warming — clear rise.<br> - Netherlands ↔ Global covariance: Dutch year-to-year temperature swings track global anomalies.<br> - CO₂ is the main driver — global carbon rise explains most warming.<br><br> RQ2 findings:<br> - GDP growth raises emissions — strong pre-1990, still significant after.<br> - Energy efficiency reduces emissions — impactful before 1990, even stronger after.<br> - Fossil-fuel share matters more post-1990 — cleaner mix = clearer emission cuts.<br>

</h2>

<h3 style="font-size: 0.8em;font-style: italic; text-align: center;">

“The world heats up, Netherlands follows - and our future now depends on smarter energy and a cleaner mix"

</h3>

# Thank you!

## Bibliography reference

<hr style="width: 1000px; margin: 10px 0; border-top: 5px solid #333;">

<h3 style="font-size: 0.4em; text-align: left;">

(1) Visual Capitalist. A global breakdown of greenhouse gas emissions by sector \[Internet\]. Vancouver (BC): Visual Capitalist; 2023 \[cited 2025 Dec 11\]. Available from: https://www.visualcapitalist.com/a-global-breakdown-of-greenhouse-gas-emissions-by-sector/
(2) Our World in Data. Average monthly surface temperature \[Internet\]. Our World in Data; 2025 \[cited 2025 Dec 11\]. Available from: https://ourworldindata.org/grapher/average-monthly-surface-temperature?tab=line&time=earliest..2025-10-15&country=NLD~Europe+(NIAID)~OWID_WRL\[1\]
(3) Our World in Data. Annual CO₂ emissions \[Internet\]. Our World in Data; 2025 \[cited 2025 Dec 11\]. Available from: https://ourworldindata.org/grapher/annual-co2-emissions-per-country?time=1960..latest&country=NLD~OWID_WRL~OWID_EUR.
(4) Our World in Data. GDP per capita - Maddison Project Database \[Internet\]. Our World in Data; 2024 \[cited 2025 Dec 11\]. Available from: https://ourworldindata.org/grapher/gdp-per-capita-maddison-project-database?tab=line&time=1960..2022&country=\~NLD.
(5) Our World in Data. GDP per capita - Maddison Project Database \[Internet\]. Our World in Data; 2024 \[cited 2025 Dec 11\]. Available from: https://ourworldindata.org/grapher/gdp-per-capita-maddison-project-database?tab=line&time=1960..2022&country=GBR~FRA~DEU&globe=1.
(6) Our World in Data. Share of primary energy consumption from fossil fuels – Using the substitution method \[Internet\]. Our World in Data; 2025 \[cited 2025 Dec 11\]. Available from: https://ourworldindata.org/grapher/fossil-fuels-share-energy?tab=line&country=NLD~OWID_EUR~OWID_WRL.
(7) Our World in Data. Primary energy consumption by source \[Internet\]. Our World in Data; 2025 \[cited 2025 Dec 11\]. Available from: https://ourworldindata.org/grapher/primary-sub-energy-source?country=\~NLD.
(8) Our World in Data. Primary energy consumption by source \[Internet\]. Our World in Data; 2025 \[cited 2025 Dec 11\]. Available from: https://ourworldindata.org/grapher/primary-sub-energy-source?country=\~FRA.
(9) Our World in Data. Primary energy consumption by source \[Internet\]. Our World in Data; 2025 \[cited 2025 Dec 11\]. Available from: https://ourworldindata.org/grapher/primary-sub-energy-source?country=\~GBR.
(10) Our World in Data. Primary energy consumption by source \[Internet\]. Our World in Data; 2025 \[cited 2025 Dec 11\]. Available from: https://ourworldindata.org/grapher/primary-sub-energy-source?country=\~DEU.
(11) von Storch H, Zwiers FW. Statistical Analysis in Climate Research. Cambridge: Cambridge University Press; 1999.
(12) Shepherd TG. Atmospheric circulation as a source of uncertainty in climate change projections. Philos Trans R Soc A Math Phys Eng Sci. 2016;374(2080):20140426. doi:10.1098/rsta.2014.0426.

</h3>
